Particular examples of planar integral point sets and their classification \footnote{ This work was carried out at Voronezh State University and supported by the Russian Science Foundation grant 19-11-00197. }

### 2025-11-17 AvdMomZvolA_examples/ips_examples.tex
[AMZA] Adjust definitions

\theoremstyle{dfn}
\newtheorem{dfn}{Definition}

@@

\begin{dfn}\label{dfn1}
	A planar integral point set (PIPS) is a set $\mathcal{P}$
	of non-collinear points in plane $\mathbb{R}^{2}$ such that
	for any pair of points $P_{1}, P_{2} \in \mathcal{P}$
	the Euclidean distance $|P_{1}P_{2}|$
	between points $P_{1}$ and $P_{2}$ is integral.
\end{dfn}

@@

\begin{dfn}
	The diameter of the integral point set $\mathcal{P}$ is defined as
	$
		\operatorname{diam(\mathcal{P})} = \underset{P_{1}, P_{2} \in
		\mathcal{P}}{\max} |P_{1}P_{2}|
		.
	$
\end{dfn}

\begin{dfn}(\cite{kemnitz1988punktmengen,kurz2005characteristic})
	The characteristic of a planar integral point set $M$ is the least positive integer $q$
	such that the area of any triangle with vertices $M_1, M_2, M_3 \in M$
	is commensurable with $\sqrt{q}$.
\end{dfn}
Characteristic of an IPS does not change if the set is moved, dilated ot flipped;
moreover, even addition or deletion of a point does not change the characteristic of IPS.

While the minimal possible diameter for planar integral point sets of given cardinality was being computed,
it was noticed~\cite{kurz2008minimum} that such diameter is attained at sets with many points on a straight line;
for some estimations on this tendency, we also refer the reader to~\cite{solymosi2003note};
(For the recently established connection between the characteristic and minimal possible diameter,
see~\cite{avdeev-lushina2025australas}.)
Thus, the following classification was introduced:
\begin{dfn}
	A PIPS $M$ is said to be in \emph{semi-general position}
	if no three points of $M$ are located in a straight line.
\end{dfn}

The most dominating examples of PIPS in semi-general position are circular sets.
\begin{dfn}
	A planar integral point sets that is situated on a circle is said to be a \textit{circular}
	point set.
\end{dfn}

So, the following constraint appeared.
\begin{dfn}
	A planar integral point set $M$ is said to be in \emph{general position}
	if no three points of $M$ are located in a straight line
	and no four points of $M$ are located in a circle.
\end{dfn}

@@

\begin{dfn}
	A planar integral point sets of $n$ points with $n-1$ points on a straight line is called
	a \textit{facher} set.
\end{dfn}
Facher sets are predominant examples of planar integral point sets.
In~\cite{antonov2008maximal}, facher sets of characteristic 1 are called \textit{semi-crabs}.
%For $9 \leq n \leq 122$, the diameter $d(2,n)$ is reached on a facher point set~\cite{kurz2008minimum}.

\begin{dfn}[\cite{avdeev2019particular}]
	A non-facher planar integral point sets situated in two parallel straight lines
	is called a \textit{rails} set.
\end{dfn}

>>>>

\newcommand{\Definition}[1]{\smallskip{\sc Definition#1.}}

@@

\Definition{ 1}\label{dfn1}
	A planar integral point set (PIPS) is a set $\mathcal{P}$
	of non-collinear points in plane $\mathbb{R}^{2}$ such that
	for any pair of points $P_{1}, P_{2} \in \mathcal{P}$
	the Euclidean distance $|P_{1}P_{2}|$
	between points $P_{1}$ and $P_{2}$ is integral.

@@

\Definition{ 2}
	The diameter of the integral point set $\mathcal{P}$ is defined as
	$
		\operatorname{diam(\mathcal{P})} = \underset{P_{1}, P_{2} \in
		\mathcal{P}}{\max} |P_{1}P_{2}|
		.
	$

\Definition{ 3}(\cite{kemnitz1988punktmengen,kurz2005characteristic})
	The characteristic of a planar integral point set $M$ is the least positive integer $q$
	such that the area of any triangle with vertices $M_1, M_2, M_3 \in M$
	is commensurable with $\sqrt{q}$.

Characteristic of an IPS does not change if the set is moved, dilated ot flipped;
moreover, even addition or deletion of a point does not change the characteristic of IPS.

While the minimal possible diameter for planar integral point sets of given cardinality was being computed,
it was noticed~\cite{kurz2008minimum} that such diameter is attained at sets with many points on a straight line;
for some estimations on this tendency, we also refer the reader to~\cite{solymosi2003note};
(For the recently established connection between the characteristic and minimal possible diameter,
see~\cite{avdeev-lushina2025australas}.)
Thus, the following classification was introduced:

\Definition{ 4}
	A PIPS $M$ is said to be in \emph{semi-general position}
	if no three points of $M$ are located in a straight line.

The most dominating examples of PIPS in semi-general position are circular sets.

\Definition{ 5}
	A planar integral point sets that is situated on a circle is said to be a \textit{circular}
	point set.

So, the following constraint appeared.

\Definition{ 6}
	A planar integral point set $M$ is said to be in \emph{general position}
	if no three points of $M$ are located in a straight line
	and no four points of $M$ are located in a circle.

@@

\Definition{ 7}
	A planar integral point sets of $n$ points with $n-1$ points on a straight line is called
	a \textit{facher} set.

Facher sets are predominant examples of planar integral point sets.
In~\cite{antonov2008maximal}, facher sets of characteristic 1 are called \textit{semi-crabs}.
%For $9 \leq n \leq 122$, the diameter $d(2,n)$ is reached on a facher point set~\cite{kurz2008minimum}.

\Definition{ 8}[\cite{avdeev2019particular}]
	A non-facher planar integral point sets situated in two parallel straight lines
	is called a \textit{rails} set.
